---
title: "Evaluation of DE analysis results on simulation data"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output: html_notebook
params:
  output_dir: NULL
editor_options:
  chunk_output_type: console
---

## 1. Preparations

```{r, setup, include=FALSE, cache=FALSE}
# Note: Do not display big number as exponential
options(scipen = 100)

library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

utl <- new.env(); source(here::here("scripts/utils/base.R"), utl); source(here::here("scripts/utils/plot.R"), utl)
h.align <- new.env(); source(here::here("scripts/helpers/align_helper.R"), h.align)
h.quant <- new.env(); source(here::here("scripts/helpers/quant_helper.R"), h.quant)
h.de <- new.env(); source(here::here("scripts/helpers/de_helper.R"), h.de)

PREFIX <- "sim"

output_dir <- ifelse(length(params$output_dir) > 0, params$output_dir, "results/figs_tbls")
input_dir <- here::here("results", c("test01_main_r"))

inputs <- bind_rows(
  data.frame(
    dataset = "sim_main",
    condition = c(0),
    input_dir = input_dir,
    stringsAsFactors = FALSE
  )
)

input_true <- here::here("results/sim_cntmat/gencode/params_sim_mod.txt")

NUM_BINS = 3
```

## 2. Load assets

```{r}

path_annotation <- here::here("share/assets/references/grch38/annotations/gencode/gencode.v31.annotation.gtf")

annotations <- utl$load_annotations()
annotation_main <- annotations$gencode

transcript_lengths <- path_annotation %>%
  utl$load_gtf(types = c("exon")) %>%
  group_by(transcript_id) %>%
  summarise(length = sum(end - start + 1)) %>%
  data.frame

exons_transcript <- path_annotation %>%
  utl$load_gtf(types = c("exon")) %>%
  group_by(transcript_id) %>%
  summarise(n_exons = n()) %>%
  data.frame

transcripts_gene <- path_annotation %>%
  utl$load_gtf(types = c("transcript")) %>%
  group_by(gene_id) %>%
  summarise(n_transcripts = n()) %>%
  data.frame

# FIXME: Gene-level biotype
features <- bind_rows(
  annotation_main %>%
    distinct(transcript_id, transcript_name, biotype) %>%
    set_names(c("feature_id", "feature_name", "biotype")),
  annotation_main %>%
    distinct(gene_id, gene_name, gene_type) %>%
    set_names(c("feature_id", "feature_name", "biotype"))
)

# DEBUG:
paths_mappability <- here::here(
  # "results/mappabilities/100",
  "results/mappabilities/100",
  c(
    "gencode.v31.transcripts.formatted.gencode.v31.transcripts.formatted.aligned.merged.sqlite"
  )
) %>% set_names(c("gencode"))

load_mappability <- function(path, feature_type) {
  conn <- RSQLite::SQLite() %>% RSQLite::dbConnect(path, synchronous = "off")

  df <- RSQLite::dbGetQuery(conn, paste0("select * from ", feature_type, "_mappabilities;")) %>%
    set_names(c("feature_id", "mappability")) %>%
    left_join(features, by = "feature_id")

  RSQLite::dbDisconnect(conn)

  df
}

feature_types <- c("gene", "transcript")

mappabilities <- paths_mappability %>%
  enframe(value = "path") %>%
  crossing(feature_type = feature_types) %>%
  mutate(data = map2(path, feature_type, ~ load_mappability(.x, .y)))

```

## 3. Load results

### 3.1 Alignment

```{r}
inputs_align <- inputs %>%
  mutate(path = map(
    input_dir,
    ~ utl$find_paths(.x, h.align$PATTERNS$transcript) %>% unlist
  )) %>%
  unnest(path) %>%
  mutate(name = paste(condition, map_chr(path, ~ .x %>% utl$to_combination()), sep = ".")) %>%
  mutate(combination = utl$.combination(name)) %>%
  utl$filter_by_combination()

results_align_metrics <- inputs_align %>%
  mutate(
    data = map(path, ~ h.align$load_result(.x, "metrics"))
  ) %>%
  mutate(mean_recall = map(data, ~ mean(.x$recall)) %>% unlist) %>%
  mutate(mean_precision = map(data, ~ mean(.x$precision)) %>% unlist) %>%
  mutate(mean_f1 = map(data, ~ mean(.x$f1)) %>% unlist)

results_align_transcript_metrics <- inputs_align %>%
  mutate(
    data = map(path, ~ h.align$load_result(.x, "transcript_metrics"))
  )

results_align_confusion_matrix <- inputs_align %>%
  mutate(
    data = map(path, ~ h.align$load_result(.x, "confusion_matrix") %>% select(-c(2)))
  )
```

### 3.2 Quantification
```{r}
result_quant_true <- input_true %>%
  h.quant$load_groundtruth() %>%
  h.quant$fill_lack(annotation_main)

result_quant_true_cpm <- result_quant_true %>%
  h.quant$to_cpm()

inputs_quant <- inputs %>%
  mutate(
    path = map(input_dir, ~ utl$find_paths(.x, h.quant$PATTERNS$transcript) %>% unlist)
  ) %>%
  unnest(path) %>%
  mutate(
    name = paste(condition, map_chr(path, ~ .x %>% utl$to_combination()), sep = ".")
  )

inputs_quant <- inputs_quant %>%
  mutate(combination = utl$.combination(name)) %>%
  filter(grepl('0.gencode-', name)) %>%
  utl$filter_by_combination()

results_quant_est <- inputs_quant %>%
  mutate(
    data = map2(
      path, name,
      ~ h.quant$load_result(.x) %>%
        h.quant$fill_lack(., annotations[[utl$annotation_used(.y)]])
    )
  ) %>%
  select(-condition, -input_dir, -path)

results_quant_est_cpm <- results_quant_est %>%
  mutate(
    data = map(data,
               ~ h.quant$to_cpm(.x) %>%
                 utl$mean_by_group(prefixes = c("CTRL", "CASE"), ordered = FALSE)
    )
  )

results_quant_est_tpm <- results_quant_est %>%
  mutate(
    data = map(data,
               ~ h.quant$to_tpm(.x, transcript_lengths) %>%
                 utl$mean_by_group(prefixes = c("CTRL", "CASE"), ordered = FALSE)
    )
  )

# CHANGED: In quantification and DE evaluations, the records have under 0.25 CPM (approximately same as 10 raw counts) values were removed to avoid inflation of the metrics.
# feature_ids_expressed <- c(list(result_quant_true), results_quant_est$data) %>%
#   map(~ .x %>% mutate(sum = rowSums(select(., where(is.numeric))))) %>%
#   map(~ .x %>% filter(sum > 0) %>% pull(feature_id)) %>%
#   Reduce(function(x, y) c(x, y), .) %>% unique

LIB_SIZE <- 40 * 10^6
.cutoff <- 10^6 / LIB_SIZE * 10
.keep <- (select(result_quant_true_cpm, where(is.numeric)) >= .cutoff) %>% apply(1, all)

feature_ids_expressed <- result_quant_true_cpm[.keep, ] %>% pull(feature_id)

feature_ids_expressed %>%
  data.frame %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "feature_ids_expressed", "data.tsv", sep = "_"))
    )

result_quant_true_cpm <- result_quant_true_cpm %>%
  filter(feature_id %in% feature_ids_expressed)

result_quant_true_cpm <- result_quant_true_cpm %>%
  mutate(mean = result_quant_true_cpm %>%
           select(where(is.numeric)) %>%
           apply(1, mean)
  )

result_quant_true_tpm <- result_quant_true %>%
  h.quant$to_tpm(transcript_lengths) %>%
  filter(feature_id %in% feature_ids_expressed)

result_quant_true_tpm <- result_quant_true_tpm %>%
  mutate(mean = result_quant_true_tpm %>%
           select(where(is.numeric)) %>%
           apply(1, mean)
  )
```

### 3.3 DE

```{r}
result_de_true <- input_true %>%
  h.de$load_groundtruth() %>%
  filter(feature_id %in% feature_ids_expressed) %>%
  h.de$fill_lack(annotation_main) %>%
  utl$set_biotype(annotation_main)

inputs_de <- inputs %>%
  mutate(path = map(input_dir, ~ utl$find_paths(.x, h.de$PATTERNS$transcript) %>% unlist)) %>%
  unnest(path) %>%
  mutate(name = paste(
    condition,
    map_chr(path, ~ .x %>% utl$to_combination()),
    sep = "."
  ))

inputs_de <- inputs_de %>%
  mutate(combination = utl$.combination(name)) %>%
  filter(grepl('0.gencode-', name)) %>%
  utl$filter_by_combination()

.load_result_de <- function(p, n) {
  p %>% message

  .f <- get(paste0("load_", utl$tool(n)), envir = h.de)
  if (utl$tool(n) == "ballgown") .f <- partial(.f, coef = -1)

  .f(p)
}

results_de_est <- inputs_de %>%
  mutate(
    data = map2(
      path, name,
      ~ .load_result_de(.x, .y) %>%
        h.de$fill_lack(., annotations[[utl$annotation_used(.y)]])
    )
  ) %>%
  select(-condition, -input_dir, -path)
```

## 4. Impact of mappability

### 4.1 Calculate metrics

```{r}
mappability_feature_ids <- mappabilities %>%
  unnest(data) %>%
  filter(feature_id %in% feature_ids_expressed) %>%
  utl$set_bin("mappability", num_bins = NUM_BINS) %>%
  select(bin, feature_id) %>%
  group_by(bin) %>%
  nest %>%
  mutate(feature_ids = map(data, ~ pull(.x, feature_id))) %>%
  select(-data) %>%
  mutate(n_features = map_int(feature_ids, length)) %>%
  arrange(bin)

# NOTE: when num_bins == 3, rename factor
if (NUM_BINS == 3) {
  mappability_feature_ids <- mappability_feature_ids %>%
    mutate(interval = bin)
  levels(mappability_feature_ids$bin) <- c("Low", "Middle", "High")
} else {
  mappability_feature_ids <- mappability_feature_ids %>%
    mutate(interval = NA)
}

mappability_metrics_align <- results_align_confusion_matrix %>%
  crossing(mappability_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.align$calc_metrics(confusion_matrix = .x, feature_ids = .y) %>%
      pivot_longer(-1, names_to = "metric")
  )) %>%
  select(dataset, name, bin, interval, n_features, metrics)

mappability_metrics_align %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_metrics_align", "data.tsv", sep = "_"))
  )

mappability_metrics_quant <- results_quant_est_cpm %>%
  crossing(mappability_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.quant$calc_metrics(est = .x, true = result_quant_true_cpm, feature_ids = .y) %>%
      pivot_longer(cols = everything(), names_to = "metric")
  )) %>%
  select(dataset, name, bin, interval, n_features, metrics)

mappability_metrics_quant %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_metrics_quant", "data.tsv", sep = "_"))
  )

mappability_metrics_de <- results_de_est %>%
  crossing(mappability_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.de$calc_metrics(est = .x, true = result_de_true, feature_ids = .y) %>%
      pivot_longer(cols = everything(), names_to = "metric")
  )) %>%
  mutate(n_tested = map2_int(
    data,
    feature_ids,
    ~ h.de$counts_n_tested(est = .x, feature_ids = .y)
  )) %>%
  select(dataset, name, bin, interval, n_features, n_tested, metrics)

mappability_metrics_de %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_metrics_de", "data.tsv", sep = "_"))
    )
```

### 4.2 Draw plots

#### 4.2.1 Alignment

```{r}
.data <- mappability_metrics_align %>%
  unnest(metrics) %>%
  group_by(dataset, metric, name, bin) %>%
  summarize(mean = mean(value)) %>%
  utl$decorate_metrics() %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

calc_coef <- function(x) {
  .x <- x %>% mutate(bin = as.numeric(factor(bin)))

  .coef <- lm(.x$value ~ .x$bin)$coefficients

  tibble(intercept = .coef[[1]],
         slope = .coef[[2]])
}

.coef <- .data %>%
  unnest(data) %>%
  select(metric, dataset, name, bin, mean) %>%
  rename(value = mean) %>%
  group_by(metric, dataset, name) %>%
  nest %>%
  mutate(coef = map(data,
                    calc_coef)) %>%
  select(-data)

.coef %>%
  unnest(coef) %>%
  write_tsv(here::here(
    output_dir,
    paste(PREFIX, "mappability_coefficients_align", "data.tsv", sep = "_")
  ))

.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = NULL,
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin",
      y = "mean",
      ylim = c(0.95, 1.00),
      xlab = "Transcript mappability",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

map2(
  .data$plot,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "mappability_line_align.pdf",
    width = 3.2 * 1.6,
    height = 3.2 * 1.5
  )
)

plots_top_align <- .data %>% select(metric, plot) %>% deframe

utl$rm_private_vars()
```

#### 4.2.2 Qunatification

```{r}
.data <- mappability_metrics_quant %>%
  unnest(metrics) %>%
  utl$decorate_metrics() %>%
  mutate(bin = factor(bin)) %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.coef <- .data %>%
  unnest(data) %>%
  select(metric, dataset, name, bin, value) %>%
  group_by(metric, dataset, name) %>%
  nest %>%
  mutate(coef = map(data,
                    calc_coef)) %>%
  select(-data)

.coef %>%
  unnest(coef) %>%
  write_tsv(here::here(
    output_dir,
    paste(PREFIX, "mappability_coefficients_quant", "data.tsv", sep = "_")
  ))

.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = NULL,
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin",
      y = "value",
      xlab = "Transcript mappability",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

map2(
  .data$plot,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "mappability_line_quant.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 3
  )
)

plots_top_quant <- .data$plot

utl$rm_private_vars()
```

#### 4.2.3 DE
```{r}
.data <- mappability_metrics_de %>%
  unnest(metrics) %>%
  utl$decorate_metrics() %>%
  mutate(bin = factor(bin)) %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.coef <- .data %>%
  unnest(data) %>%
  select(metric, dataset, name, bin, value) %>%
  group_by(metric, dataset, name) %>%
  nest %>%
  mutate(coef = map(data,
                    calc_coef)) %>%
  select(-data)

.coef %>%
  unnest(coef) %>%
  write_tsv(here::here(
    output_dir,
    paste(PREFIX, "mappability_coefficients_de", "data.tsv", sep = "_")
  ))

.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = NULL,
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin",
      y = "value",
      xlab = "Transcript mappability",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

map2(
  .data$plot,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "mappability_line_de.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 3
  )
)

plots_top_de <- .data$plot

utl$rm_private_vars()
```

## 5. Impact of transcript abundance

### 5.1 Calculate metrics

```{r}
# NOTE: bin, ctrl: feature_ids1, case: feature_ids2
abundance_feature_ids <- result_quant_true_cpm %>%
  filter(feature_id %in% feature_ids_expressed) %>%
  select(-mean) %>%
  pivot_longer(-feature_id, names_to = "group", values_to = "value") %>%
  utl$set_bin("value", num_bins = NUM_BINS, exclude = FALSE) %>%
  select(feature_id, group, bin) %>%
  pivot_wider(names_from = group, values_from = feature_id) %>%
  rename(feature_ids1 = "ctrl", feature_ids2 = "case") %>%
  mutate(n_features1 = map_int(feature_ids1, length)) %>%
  mutate(n_features2 = map_int(feature_ids2, length)) %>%
  arrange(bin)

# NOTE: when num_bins == 3, rename factor
if (NUM_BINS == 3) {
  abundance_feature_ids <- abundance_feature_ids %>%
    mutate(interval = bin)
  levels(abundance_feature_ids$bin) <- c("Low", "Middle", "High")
} else {
  abundance_feature_ids <- abundance_feature_ids %>%
    mutate(interval = bin)
}

# NOTE: for evaluation of DE
mean_abundance_feature_ids <- result_quant_true_cpm %>%
  filter(feature_id %in% feature_ids_expressed) %>%
  select(feature_id, mean) %>%
  utl$set_bin("mean", num_bins = NUM_BINS, exclude = FALSE) %>%
  select(bin, feature_id) %>%
  group_by(bin) %>%
  nest %>%
  mutate(feature_ids = map(data, ~ pull(.x, feature_id))) %>%
  mutate(n_features = map_int(feature_ids, length)) %>%
  select(-data) %>%
  arrange(bin)

# NOTE: when num_bins == 3, rename factor
if (NUM_BINS == 3) {
  mean_abundance_feature_ids <- mean_abundance_feature_ids %>%
    mutate(interval = bin)
  levels(mean_abundance_feature_ids$bin) <- c("Low", "Middle", "High")
} else {
  mean_abundance_feature_ids <- mean_abundance_feature_ids %>%
    mutate(interval = NA)
}

abundance_metrics_align <- results_align_confusion_matrix %>%
  crossing(mean_abundance_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.align$calc_metrics(confusion_matrix = .x, feature_ids = .y) %>%
      pivot_longer(-1, names_to = "metric")
  )) %>%
  select(dataset, name, bin, interval, n_features, metrics)

abundance_metrics_align %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "abundance_metrics_align", "data.tsv", sep = "_"))
  )

.calc_metrics2 <- partial(h.quant$calc_metrics2, true = result_quant_true_cpm)

abundance_metrics_quant <- results_quant_est_cpm %>%
  crossing(filter(abundance_feature_ids, !is.na(bin))) %>%
  mutate(metrics = pmap(
    list(
      est = data,
      feature_ids1 = feature_ids1,
      feature_ids2 = feature_ids2
    ),
    .calc_metrics2
  ) %>% map( ~ .x %>% pivot_longer(cols = everything(), names_to = "metric"))) %>%
  select(dataset, name, bin, interval, n_features1, n_features2, metrics)

abundance_metrics_quant %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "abundance_metrics_quant", "data.tsv", sep = "_"))
    )

abundance_metrics_de <- results_de_est %>%
  crossing(filter(mean_abundance_feature_ids, !is.na(bin))) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.de$calc_metrics(est = .x, true = result_de_true, feature_ids = .y) %>%
      pivot_longer(cols = everything(), names_to = "metric")
  )) %>%
  mutate(n_tested = map2_int(
    data,
    feature_ids,
    ~ h.de$counts_n_tested(est = .x, feature_ids = .y)
  )) %>%
  select(dataset, name, bin, interval, n_features, n_tested, metrics)

abundance_metrics_de %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "abundance_metrics_de", "data.tsv", sep = "_"))
    )
```

### 5.2 Draw line plot

#### 5.2.1 Alignment

```{r}
.data <- abundance_metrics_align %>%
  unnest(metrics) %>%
  group_by(dataset, metric, name, bin) %>%
  summarize(mean = mean(value)) %>%
  utl$decorate_metrics() %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.coef <- .data %>%
  unnest(data) %>%
  select(metric, dataset, name, bin, mean) %>%
  rename(value = mean) %>%
  group_by(metric, dataset, name) %>%
  nest %>%
  mutate(coef = map(data,
                    calc_coef)) %>%
  select(-data)

.coef %>%
  unnest(coef) %>%
  write_tsv(here::here(
    output_dir,
    paste(PREFIX, "abundance_coefficients_align", "data.tsv", sep = "_")
  ))

.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = NULL,
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin",
      y = "mean",
      ylim = c(0.95, 1.00),
      xlab = "Transcript abundance",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

map2(
  .data$plot,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "abundance_line_align.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 3
  )
)

plots_right_align <- .data$plot

utl$rm_private_vars()
```

#### 5.2.2 Qunatification

```{r}
.data <- abundance_metrics_quant %>%
  unnest(metrics) %>%
  utl$decorate_metrics() %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.coef <- .data %>%
  unnest(data) %>%
  select(metric, dataset, name, bin, value) %>%
  group_by(metric, dataset, name) %>%
  nest %>%
  mutate(coef = map(data,
                    calc_coef)) %>%
  select(-data)

.coef %>%
  unnest(coef) %>%
  write_tsv(here::here(
    output_dir,
    paste(PREFIX, "abundance_coefficients_quant", "data.tsv", sep = "_")
  ))

.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = NULL,
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin",
      y = "value",
      xlab = "Transcript abundance",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

map2(
  .data$plot,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "abundance_line_quant.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 3
  )
)

plots_right_quant <- .data$plot

utl$rm_private_vars()
```

#### 5.2.3 DE
```{r}
.data <- abundance_metrics_de %>%
  unnest(metrics) %>%
  utl$decorate_metrics() %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.coef <- .data %>%
  unnest(data) %>%
  select(metric, dataset, name, bin, value) %>%
  group_by(metric, dataset, name) %>%
  nest %>%
  mutate(coef = map(data,
                    calc_coef)) %>%
  select(-data)

.coef %>%
  unnest(coef) %>%
  write_tsv(here::here(
    output_dir,
    paste(PREFIX, "abundance_coefficients_de", "data.tsv", sep = "_")
  ))

.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = NULL,
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin",
      y = "value",
      xlab = "Transcript abundance",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

map2(
  .data$plot,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "abundance_line_de.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 3
  )
)

plots_right_de <- .data$plot

utl$rm_private_vars()
```

## 6. Impact of mappability and abundance

### 6.1 Calculate metrics

```{r}
mappability_abundance_feature_ids1 <- left_join(
  mappability_feature_ids %>%
    unnest(feature_ids),
  abundance_feature_ids %>%
    select(-feature_ids2) %>%
    unnest(feature_ids1) %>%
    dplyr::rename(feature_ids = "feature_ids1"),
  by = "feature_ids") %>%
  dplyr::rename(feature_id = "feature_ids") %>%
  relocate(feature_id) %>%
  arrange(bin.x, bin.y)

mappability_abundance_feature_ids2 <- left_join(
  mappability_feature_ids %>%
    unnest(feature_ids),
  abundance_feature_ids %>%
    select(-feature_ids1) %>%
    unnest(feature_ids2) %>%
    dplyr::rename(feature_ids = "feature_ids2"),
  by = "feature_ids") %>%
  dplyr::rename(feature_id = "feature_ids") %>%
  relocate(feature_id) %>%
  arrange(bin.x, bin.y)

# NOTE: bin.x for mappability; bin.y for abundance
# NOTE: bin.y including NA value derived from undefined records
mappability_abundance_feature_ids <- left_join(
  mappability_abundance_feature_ids1 %>%
    group_by(bin.x, interval.x, bin.y, interval.y) %>%
    nest,
  mappability_abundance_feature_ids2 %>%
    group_by(bin.x, interval.x, bin.y, interval.y) %>%
    nest,
  by = c("bin.x", "bin.y", "interval.x", "interval.y")
) %>%
  mutate(feature_ids1 = map(data.x, ~ pull(.x, feature_id))) %>%
  mutate(feature_ids2 = map(data.y, ~ pull(.x, feature_id))) %>%
  select(!starts_with("data"))

mappability_abundance_feature_ids_count_matrix <- cbind(
  mappability_abundance_feature_ids1 %>%
    group_by(bin.x, bin.y) %>%
    summarise(n = n()) %>%
    pivot_wider(everything(), names_from = bin.y, values_from = n) %>%
    rename_if(is.numeric, ~ paste0("ctrl:", .x)) %>%
    ungroup,
  mappability_abundance_feature_ids2 %>%
    group_by(bin.x, bin.y) %>%
    summarise(n = n()) %>%
    pivot_wider(everything(), names_from = bin.y, values_from = n) %>%
    rename_if(is.numeric, ~ paste0("case:", .x)) %>%
    ungroup %>%
    select(-1)
)

mappability_abundance_feature_ids_count_matrix %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_abundance_feature_ids_count_matrix", "data.tsv", sep = "_"))
    )

# NOTE: bin.x for mappability; bin.y for abundance
# NOTE: bin.y including NA value derived from undefined records
mappability_mean_abundance_feature_ids <- left_join(
  mappability_feature_ids %>%
    unnest(feature_ids),
  mean_abundance_feature_ids %>%
    unnest(feature_ids),
  by = "feature_ids") %>%
  rename(feature_id = "feature_ids") %>%
  relocate(feature_id) %>%
  arrange(bin.x, bin.y) %>%
  group_by(bin.x, bin.y, interval.x, interval.y) %>%
  nest %>%
  mutate(feature_ids = map(
    data,
    ~ pull(.x, feature_id)
  )) %>%
  select(-data) %>%
  mutate(n_features = map_int(feature_ids, length))

mappability_mean_abundance_feature_ids_count_matrix <- cbind(
  mappability_mean_abundance_feature_ids %>%
    ungroup %>%
    select(1:4),
  mappability_mean_abundance_feature_ids %>%
    ungroup %>%
    select(-(1:4)) %>%
    apply(c(1, 2), function(x) length(x[[1]]))
)

mappability_mean_abundance_feature_ids_count_matrix %>%
  rename_with(~ gsub(".y", ".abundance", gsub(".x", ".mappability", .x, fixed = TRUE), fixed = TRUE)) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_abundance_feature_ids_count_matrix", "data.tsv", sep = "_"))
    )

# NOTE: UNUSED
mappability_abundance_metrics_align <- results_align_confusion_matrix %>%
  crossing(filter(mappability_mean_abundance_feature_ids, !is.na(bin.y))) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.align$calc_metrics(confusion_matrix = .x, feature_ids = .y) %>%
      pivot_longer(cols = -1, names_to = "metric")
  )) %>%
  select(dataset, name, bin.x, interval.x, , bin.y, interval.y, n_features, metrics)

mappability_abundance_metrics_align %>%
  rename_with(~ gsub(".y", ".abundance", gsub(".x", ".mappability", .x, fixed = TRUE), fixed = TRUE)) %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_abundance_metrics_align", "data.tsv", sep = "_"))
    )

.calc_metrics2 <- partial(h.quant$calc_metrics2, true = result_quant_true_cpm)

mappability_abundance_metrics_quant <- results_quant_est_cpm %>%
  crossing(filter(mappability_abundance_feature_ids, !is.na(bin.y))) %>%
  mutate(metrics = pmap(
    list(
      est = data,
      feature_ids1 = feature_ids1,
      feature_ids2 = feature_ids2
    ),
    .calc_metrics2
  ) %>% map( ~ .x %>% pivot_longer(cols = everything(), names_to = "metric"))) %>%
  mutate(n_features1 = map_int(feature_ids1, length)) %>%
  mutate(n_features2 = map_int(feature_ids2, length)) %>%
  select(dataset, name, bin.x, interval.x, bin.y, interval.y, n_features1, n_features2, metrics)

mappability_abundance_metrics_quant %>%
  rename_with(~ gsub(".y", ".abundance", gsub(".x", ".mappability", .x, fixed = TRUE), fixed = TRUE)) %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_abundance_metrics_quant", "data.tsv", sep = "_"))
    )

mappability_abundance_metrics_de <- results_de_est %>%
  crossing(filter(mappability_mean_abundance_feature_ids, !is.na(bin.y))) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.de$calc_metrics(est = .x, true = result_de_true, feature_ids = .y) %>%
      pivot_longer(cols = everything(), names_to = "metric")
  )) %>%
  mutate(n_features = map_int(feature_ids, length)) %>%
  mutate(n_tested = map2_int(
    data,
    feature_ids,
    ~ h.de$counts_n_tested(est = .x, feature_ids = .y)
  )) %>%
  select(dataset, name, bin.x, interval.x, bin.y, interval.y, n_features, n_tested, metrics)

mappability_abundance_metrics_de %>%
  rename_with(~ gsub(".y", ".abundance", gsub(".x", ".mappability", .x, fixed = TRUE), fixed = TRUE)) %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_abundance_metrics_de", "data.tsv", sep = "_"))
    )
```

### 6.2 Draw plots

#### 6.2.0 Common definition
```{r}
get_legend_intervals <- function(data_, metric, var) {
  n_fills <- data_[[var]] %>% unique %>% length

  g <- ggpubr::ggbarplot(data_, "bin.x", "bin.y", fill = var)
  g <- g + scale_fill_manual(values = rep("white", n_fills))
  g <- g + theme(
    legend.key = element_blank(),
    legend.key.size = unit(0, "pt")
  )

  ggpubr::get_legend(g)
}

annotate_ext_legend <- function(g, l) {
 g %>%
    ggpubr::annotate_figure(bottom = l)
}

```

#### 6.2.1 Alignment

```{r}
.data <- mappability_abundance_metrics_align %>%
  unnest(metrics) %>%
  group_by(dataset, metric, name, bin.x, interval.x, bin.y, interval.y) %>%
  summarize(mean = mean(value)) %>%
  utl$decorate_metrics() %>%
  mutate(label_bin.y = fct_relabel(bin.y, ~ paste0("True abundance: ", .x))) %>%
  mutate(label_bin.x = fct_relabel(bin.x, ~ paste0("Transcript mappablity: ", .x))) %>%
  mutate(`Intervals of transcript mappability:` = factor(paste0(bin.x, "; ", interval.x))) %>%
  mutate(`Intervals of true abundance:` = factor(paste0(bin.y, "; ", interval.y))) %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = c("label_bin.y"),
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin.x",
      y = "mean",
      xlab = "Transcript mappability",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

.data$legend1 <- .data %>%
  select(data, metric) %>%
  mutate(var = "Intervals of transcript mappability:") %>%
  pmap(get_legend_intervals)

.data$legend2 <- .data %>%
  select(data, metric) %>%
  mutate(var = "Intervals of true abundance:") %>%
  pmap(get_legend_intervals)

.data <- .data %>%
  mutate(plot_annotated = map2(plot, legend1, annotate_ext_legend)) %>%
  mutate(plot_annotated = map2(plot_annotated, legend2, annotate_ext_legend))

map2(
  .data$plot_annotated,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "mappability_abundance_line_align.pdf",
    width = 3.2 * 1.6,
    height = 3.2 * 1.6
  )
)

plots_main_align <- .data$plot

# NOTE: facet by mappability
.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = c("label_bin.x"),
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin.y",
      y = "mean",
      xlab = "True transcript abundance",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot2 <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

utl$rm_private_vars()
```

#### 6.2.2 Qunatification

```{r}
.data <- mappability_abundance_metrics_quant %>%
  unnest(metrics) %>%
  utl$decorate_metrics() %>%
  mutate(label_bin.y = fct_relabel(bin.y, ~ paste0("True abundance: ", .x))) %>%
  mutate(`Intervals of transcript mappability:` = factor(paste0(bin.x, "; ", interval.x))) %>%
  mutate(`Intervals of true abundance:` = factor(paste0(bin.y, "; ", interval.y))) %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = c("label_bin.y"),
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin.x",
      y = "value",
      xlab = "Transcript mappability",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

.data$legend1 <- .data %>%
  select(data, metric) %>%
  mutate(var = "Intervals of transcript mappability:") %>%
  pmap(get_legend_intervals)

.data$legend2 <- .data %>%
  select(data, metric) %>%
  mutate(var = "Intervals of true abundance:") %>%
  pmap(get_legend_intervals)

.data <- .data %>%
  mutate(plot_annotated = map2(plot, legend1, annotate_ext_legend)) %>%
  mutate(plot_annotated = map2(plot_annotated, legend2, annotate_ext_legend))

map2(
  .data$plot_annotated,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "mappability_abundance_line_quant.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 1.6
  )
)

plots_main_quant <- .data$plot
plots_main_quant_annotated <- .data$plot_annotated

utl$rm_private_vars()
```

#### 6.2.3 DE
```{r}
.data <- mappability_abundance_metrics_de %>%
  unnest(metrics) %>%
  utl$decorate_metrics() %>%
  mutate(label_bin.y = fct_relabel(bin.y, ~ paste0("True abundance: ", .x))) %>%
  mutate(`Intervals of transcript mappability:` = factor(paste0(bin.x, "; ", interval.x))) %>%
  mutate(`Intervals of true abundance:` = factor(paste0(bin.y, "; ", interval.y))) %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.draw <- function(data_, metric) {
  g <- utl$plot_line(
    data = data_,
    group = "abbr",
    facet.by = c("label_bin.y"),
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "bin.x",
      y = "value",
      xlab = "Transcript mappability",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

.data$legend1 <- .data %>%
  select(data, metric) %>%
  mutate(var = "Intervals of transcript mappability:") %>%
  pmap(get_legend_intervals)

.data$legend2 <- .data %>%
  select(data, metric) %>%
  mutate(var = "Intervals of true abundance:") %>%
  pmap(get_legend_intervals)

.data <- .data %>%
  mutate(plot_annotated = map2(plot, legend1, annotate_ext_legend)) %>%
  mutate(plot_annotated = map2(plot_annotated, legend2, annotate_ext_legend))

map2(
  .data$plot_annotated,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "mappability_abundance_line_de.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 1.6
  )
)

plots_main_de <- .data$plot

utl$rm_private_vars()
```

## 7. Arrange plots (align and quant)

```{r}
.add_lab <- function(g, text) {
  (g + theme(plot.margin= unit(c(2, 1.5, 0.5, 0.5), "lines"))) %>%
    ggpubr::annotate_figure(fig.lab = text, fig.lab.size = 21, fig.lab.face = "bold")
}

g <- ggpubr::ggarrange(
  ggpubr::ggarrange(plots_top_align$f1 %>% .add_lab("(A)"),
                    NULL,
                    ncol = 2, widths = c(1.1, 0.90)),
  plots_main_quant_annotated[[1]] %>% .add_lab("(B)"),
  ncol = 1,
  nrow = 2,
  align = "hv",
  heights = c(0.9, 1.1),
  common.legend = FALSE,
  legend = "bottom"
)

g %>% utl$save_plot("f1_spearman", "arranged_align_quant.pdf", width = 3.2 * 3.2, height = 3.2 * 3.2)

utl$rm_private_vars()
```

## 8. Additional: Characterization of biotypes

### 8.1 Calculate metrics

#### Prepares
```{r}
biotype_feature_ids <- bind_rows(
  annotation_main %>%
    distinct(feature_id, biotype) %>%
    utl$split_tibble("biotype") %>%
    map(~ intersect(.$feature_id, feature_ids_expressed)) %>%
    enframe(name = "biotype", value = "feature_ids"),
  annotation_main %>%
    mutate(biotype = "all") %>%
    distinct(feature_id, biotype) %>%
    utl$split_tibble("biotype") %>%
    map(~ intersect(.$feature_id, feature_ids_expressed)) %>%
    enframe(name = "biotype", value = "feature_ids")
) %>% mutate(n_features = map_int(feature_ids, length))

.biotype_feature_ids_unnested <- biotype_feature_ids %>%
  unnest(feature_ids) %>%
  rename(feature_id = feature_ids)

# NOTE: bin.x for mappability; bin.y for abundance
mappability_abundance_biotype_feature_ids <- left_join(
  mappability_abundance_feature_ids1 %>%
    left_join(.biotype_feature_ids_unnested, by = "feature_id") %>%
    group_by(bin.x, interval.x, bin.y, interval.y, biotype) %>%
    nest,
  mappability_abundance_feature_ids2 %>%
    left_join(.biotype_feature_ids_unnested, by = "feature_id") %>%
    group_by(bin.x, interval.x, bin.y, interval.y, biotype) %>%
    nest,
  by = c("bin.x", "bin.y", "interval.x", "interval.y", "biotype")
) %>%
  mutate(feature_ids1 = map(data.x, ~ pull(.x, feature_id))) %>%
  mutate(feature_ids2 = map(data.y, ~ pull(.x, feature_id))) %>%
  select(!starts_with("data")) %>%
  mutate(n_features1 = map_int(feature_ids1, length)) %>%
  mutate(n_features2 = map_int(feature_ids2, length)) %>%
  arrange(bin.x, bin.y, biotype)

abundance_biotype_feature_ids <- left_join(
  mappability_abundance_feature_ids1 %>%
    left_join(.biotype_feature_ids_unnested, by = "feature_id") %>%
    select(-bin.x, -interval.x) %>%
    group_by(bin.y, interval.y, biotype) %>%
    nest,
  mappability_abundance_feature_ids2 %>%
    left_join(.biotype_feature_ids_unnested, by = "feature_id") %>%
    select(-bin.x, -interval.x) %>%
    group_by(bin.y, interval.y, biotype) %>%
    nest,
  by = c("bin.y", "interval.y", "biotype")
) %>%
  mutate(feature_ids1 = map(data.x, ~ pull(.x, feature_id))) %>%
  mutate(feature_ids2 = map(data.y, ~ pull(.x, feature_id))) %>%
  select(!starts_with("data")) %>%
  mutate(n_features1 = map_int(feature_ids1, length)) %>%
  mutate(n_features2 = map_int(feature_ids2, length)) %>%
  arrange(bin.y, biotype)

mappability_mean_abundance_biotype_feature_ids <- mappability_mean_abundance_feature_ids %>%
  select(bin.x, interval.x, bin.y, interval.y, feature_ids) %>%
  unnest(feature_ids) %>%
  rename(feature_id = feature_ids) %>%
  left_join(.biotype_feature_ids_unnested, by = "feature_id") %>%
  group_by(bin.x, interval.x, bin.y, interval.y, biotype) %>%
  nest %>%
  mutate(feature_ids = map(data, ~ pull(.x, feature_id))) %>%
  select(!starts_with("data")) %>%
  mutate(n_features = map_int(feature_ids, length)) %>%
  arrange(bin.x, bin.y, biotype)

mappability_biotype_feature_ids <- mappability_mean_abundance_biotype_feature_ids %>%
  ungroup %>%
  select(-bin.y, -interval.y) %>%
  unnest(feature_ids) %>%
  rename(feature_id = feature_ids) %>%
  group_by(bin.x, interval.x, biotype) %>%
  nest %>%
  mutate(feature_ids = map(data, ~ pull(.x, feature_id))) %>%
  select(!starts_with("data")) %>%
  mutate(n_features = map_int(feature_ids, length)) %>%
  arrange(bin.x, biotype)

mean_abundance_biotype_feature_ids <- mappability_mean_abundance_biotype_feature_ids %>%
  ungroup %>%
  select(-bin.x, -interval.x) %>%
  unnest(feature_ids) %>%
  rename(feature_id = feature_ids) %>%
  group_by(bin.y, interval.y, biotype) %>%
  nest %>%
  mutate(feature_ids = map(data, ~ pull(.x, feature_id))) %>%
  select(!starts_with("data")) %>%
  mutate(n_features = map_int(feature_ids, length)) %>%
  arrange(bin.y, biotype)
```

#### For each biotype
```{r}
biotype_metrics_align <- results_align_confusion_matrix %>%
  crossing(biotype_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.align$calc_metrics(confusion_matrix = .x, feature_ids = .y) %>%
      pivot_longer(-1, names_to = "metric")
  )) %>%
  select(dataset, name, biotype, n_features, metrics)

biotype_metrics_align %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "biotype_metrics_align", "data.tsv", sep = "_"))
  )

biotype_metrics_quant <- results_quant_est_cpm %>%
  crossing(biotype_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.quant$calc_metrics(est = .x, true = result_quant_true_cpm, feature_ids = .y) %>%
      pivot_longer(cols = everything(), names_to = "metric")
  )) %>%
  select(dataset, name, biotype, n_features, metrics)

biotype_metrics_quant %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "biotype_metrics_quant", "data.tsv", sep = "_"))
  )

biotype_metrics_de <- results_de_est %>%
  crossing(biotype_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.de$calc_metrics(est = .x, true = result_de_true, feature_ids = .y) %>%
      pivot_longer(cols = everything(), names_to = "metric")
  )) %>%
  mutate(n_tested = map2_int(
    data,
    feature_ids,
    ~ h.de$counts_n_tested(est = .x, feature_ids = .y)
  )) %>%
  select(dataset, name, biotype, n_features, n_tested, metrics)

biotype_metrics_de %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "biotype_metrics_de", "data.tsv", sep = "_"))
    )
```

#### For each biotype - mappability
```{r}
mappability_biotype_metrics_align <- results_align_confusion_matrix %>%
  crossing(mappability_biotype_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.align$calc_metrics(confusion_matrix = .x, feature_ids = .y) %>%
      pivot_longer(-1, names_to = "metric")
  )) %>%
  select(dataset, name, bin.x, interval.x, biotype, n_features, metrics)

mappability_biotype_metrics_align %>%
  rename_with(~ gsub(".y", ".abundance", gsub(".x", ".mappability", .x, fixed = TRUE), fixed = TRUE)) %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_biotype_metrics_align", "data.tsv", sep = "_"))
  )

.calc_metrics2 <- partial(h.quant$calc_metrics2, true = result_quant_true_cpm)

mappability_abundance_biotype_metrics_quant <- results_quant_est_cpm %>%
  crossing(filter(mappability_abundance_biotype_feature_ids, !is.na(bin.y))) %>%
  mutate(metrics = pmap(
    list(
      est = data,
      feature_ids1 = feature_ids1,
      feature_ids2 = feature_ids2
    ),
    .calc_metrics2
  ) %>% map( ~ .x %>% pivot_longer(cols = everything(), names_to = "metric"))) %>%
  select(dataset, name, bin.x, interval.x, bin.y, interval.y, biotype, n_features1, n_features2, metrics)

mappability_abundance_biotype_metrics_quant %>%
  rename_with(~ gsub(".y", ".abundance", gsub(".x", ".mappability", .x, fixed = TRUE), fixed = TRUE)) %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_abundance_biotype_metrics_quant", "data.tsv", sep = "_"))
  )

mappability_mean_abundance_biotype_metrics_de <- results_de_est %>%
  crossing(mappability_mean_abundance_biotype_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.de$calc_metrics(est = .x, true = result_de_true, feature_ids = .y) %>%
      pivot_longer(cols = everything(), names_to = "metric")
  )) %>%
  mutate(n_tested = map2_int(
    data,
    feature_ids,
    ~ h.de$counts_n_tested(est = .x, feature_ids = .y)
  )) %>%
  select(dataset, name, bin.x, interval.x, bin.y, interval.y, biotype, n_features, n_tested, metrics)

mappability_mean_abundance_biotype_metrics_de %>%
  rename_with(~ gsub(".y", ".abundance", gsub(".x", ".mappability", .x, fixed = TRUE), fixed = TRUE)) %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "mappability_abundance_biotype_metrics_de", "data.tsv", sep = "_"))
    )
```

#### For each abundance - biotype
```{r}
.calc_metrics2 <- partial(h.quant$calc_metrics2, true = result_quant_true_cpm)

abundance_biotype_metrics_quant <- results_quant_est_cpm %>%
  crossing(filter(abundance_biotype_feature_ids, !is.na(bin.y))) %>%
  mutate(metrics = pmap(
    list(
      est = data,
      feature_ids1 = feature_ids1,
      feature_ids2 = feature_ids2
    ),
    .calc_metrics2
  ) %>% map( ~ .x %>% pivot_longer(cols = everything(), names_to = "metric"))) %>%
  select(dataset, name, bin.y, interval.y, biotype, n_features1, n_features2, metrics)

abundance_biotype_metrics_quant %>%
  rename_with(~ gsub(".y", ".abundance", gsub(".x", ".mappability", .x, fixed = TRUE), fixed = TRUE)) %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "abundance_biotype_metrics_quant", "data.tsv", sep = "_"))
  )

mean_abundance_biotype_metrics_de <- results_de_est %>%
  crossing(mean_abundance_biotype_feature_ids) %>%
  mutate(metrics = map2(
    data,
    feature_ids,
    ~ h.de$calc_metrics(est = .x, true = result_de_true, feature_ids = .y) %>%
      pivot_longer(cols = everything(), names_to = "metric")
  )) %>%
  mutate(n_tested = map2_int(
    data,
    feature_ids,
    ~ h.de$counts_n_tested(est = .x, feature_ids = .y)
  )) %>%
  select(dataset, name, bin.y, interval.y, biotype, n_features, n_tested, metrics)

mean_abundance_biotype_metrics_de %>%
  rename_with(~ gsub(".y", ".abundance", gsub(".x", ".mappability", .x, fixed = TRUE), fixed = TRUE)) %>%
  unnest(metrics) %>%
  write_tsv(
    here::here(output_dir, paste(PREFIX, "abundance_biotype_metrics_de", "data.tsv", sep = "_"))
    )
```

### 8.2 Draw plots

#### 8.2.1 Alignment
```{r}
.data <- biotype_metrics_align %>%
  unnest(metrics) %>%
  group_by(dataset, metric, name, biotype) %>%
  summarize(mean = mean(value)) %>%
  utl$decorate_metrics() %>%
  utl$filter_by_combination("main") %>%
  mutate(biotype = utl$to_factor(biotype)) %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.draw <- function(data_, metric) {
  g <- utl$plot_bar(
    data = data_,
    group = "abbr",
    facet.by = NULL,
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "biotype",
      y = "mean",
      ylim = c(0.925, 1.00),
      xlab = "",
      ylab = utl$to_lab(metric),
      add = "mean"
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

map2(
  .data$plot,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "biotype_bar_align.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 3
  )
)

utl$rm_private_vars()
```

#### 8.2.2 Quantification

```{r}
.data <- biotype_metrics_quant %>%
  unnest(metrics) %>%
  group_by(dataset, metric, name, biotype) %>%
  summarize(mean = mean(value)) %>%
  utl$decorate_metrics() %>%
  utl$filter_by_combination("main") %>%
  mutate(biotype = utl$to_factor(biotype)) %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.draw <- function(data_, metric) {
  g <- utl$plot_bar(
    data = data_,
    group = "abbr",
    facet.by = NULL,
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "biotype",
      y = "mean",
      xlab = "",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

map2(
  .data$plot,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "biotype_bar_quant.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 3
  )
)

utl$rm_private_vars()
```

#### 8.2.3 DE

```{r}
.data <- biotype_metrics_de %>%
  unnest(metrics) %>%
  group_by(dataset, metric, name, biotype) %>%
  summarize(mean = mean(value)) %>%
  utl$decorate_metrics() %>%
  utl$filter_by_combination("main") %>%
  mutate(biotype = utl$to_factor(biotype)) %>%
  group_by(dataset, metric) %>%
  nest %>%
  ungroup(dataset)

.draw <- function(data_, metric) {
  g <- utl$plot_bar(
    data = data_,
    group = "abbr",
    facet.by = NULL,
    tilt = FALSE,
    theme_ = NULL,
    list(
      x = "biotype",
      y = "mean",
      xlab = "",
      ylab = utl$to_lab(metric)
    )
  )
}

.data$plot <- .data %>%
  select(data, metric) %>%
  pmap(.draw)

map2(
  .data$plot,
  .data$metric,
  ~ .x %>% utl$save_plot(
    .y,
    "biotype_bar_de.pdf",
    width = 3.2 * 3.2,
    height = 3.2 * 3
  )
)

utl$rm_private_vars()
```

